Constraining Quiescent Supermassive Black Hole Environments.

### 2014-09-11 ms.tex
Ang momentum section and other improvements in language

\documentclass[usenatbib,fleqn]{mn2e}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{outdir=./figures/}
\graphicspath{{./figures/}}
\usepackage{url}
\usepackage{aas_macros}
\usepackage{astro}
%\usepackage{natbib}

@@

\newcommand{\Menc}{M_{\rm enc}}
\newcommand{\rhostar}{\rho_*}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\Mseight}{M_{\star,8}}
\newcommand{\Mbh}[1][]{M_{\bullet#1}}
\newcommand{\Mbheight}{M_{\bullet,8}}
\newcommand{\MbhNorm}{\frac{\Mbh}{10^8 \Msun}}

@@

%%Maybe active galaxies constitute 10% need reference here?
Supermassive black holes (SMBHs) are present in the centers of most,
if not all nearby galaxies (see reviews by,
e.g. \citealt{KormendyRichstone:1995a};
\citealt{FerrareseFord:2005a}). Roughly $\sim 1\%$ of these manifest
themselves as luminous Active Galactic Nuclei. However, quiescent
black holes constitute a silent majority.

In order to understand why most galaxies appear to be inactive, it is
necessary to understand the gas environment of black holes, as the gas
density around the black hole sets its accretion rate.  The accretion
rate also has implications for rate of black hole growth.
%AG:Perhaps need more detail here.

Constraints on the SMBH accretion rates at the low end of the SMBH
mass function also have implications for the black occupation
fraction.  For example, \citealt{MillerGallo+:2014a} used an observed
relation between $\Mbh$ and nuclear x-ray luminosity $L_x$ to
constrain the BH occupation fraction.  However, they extrapolate
(assuming a single power law) their observed relationship below their
detection threshold.  However, it is possible that the $L_x$-$\Mbh$
relationship steepens at smaller $\Mbh$, and thus the black hole
occupation fraction could be underestimated at low $\Mbh$.
%AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

@@

However, this approach has limitations. Due to resolution issues it is
almost never possible to observationally determine the gas density at
the black sphere of influence, and such studies are forced to
extrapolate observed gas temperatures and densities to smaller scales.
In fact, the temperature profile is assumed to be flat inside of the
Bondi radius.  In reality, the cusp in the velocity dispersion near
the SMBH, should cause a cusp in the gas temperature profile (assuming
the kinetic energy of stellar winds is efficiently themalized in
shocks).
%%AG-give sense of scale.

@@

We compare our calculated gas profiles to those inferred from Chandra
X-ray observations by \citealt{AllenDunn+:2006a} and
\citealt{RussellMcNamara+:2013a}. Additionally, we chack how well our
results are able to reproduced relationships between $L_x$ and $\Mbh$
in \citealt{MillerGallo+:2014a}. In performing these comparisons we
hope to determine how well the gas environments of quiescent galaxies
may be understood using the relatively well understood physics of
stellar winds and supernova heating.
 %%AG--say something about x-ray observations.

@@

This is a broken power law which transitions from an inner power law
slope, $\gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  The corresponding (spherically symmetric) stellar
density, $\rhostar\sim r^{-1-\gamma}$ for $r \ll \rb$ and
$\rhostar\sim r^{-1-\beta}$ for $r \ll \rb$.  $\gamma\simeq1$ for cusp
galaxies, while $\gamma\simeq0$ for core galaxies.
%AG:Is there a systematic dependence of beta on galaxy type$.

@@

\section{Model}
\label{sec:model}
We use the same model as \citealt{Quataert:2004a} to find a steady
state gas density near the SMBH for each of the galaxies within our
sample.  Gas is supplied by winds from the stellar population in the
galactic center. Assuming a rate of mass injection, $q$, spherical
symmetry and an ideal gas equation of state.

\begin{align}
&\drhodt+\frac{1}{r^2}\dxdy{\rho r^2 v}{r}=q\\
&\rho \frac{dv}{dt}=-\dpdr-\rho \dphidr-q v\\
&\rho T \frac{ds}{dt}=q\left[\ke+\kew+\gammaf \cs \right]
\end{align}

@@

$q$ represent the rate of injection of gas mass per unit volume, and
is derived from the stellar density, $\rho_*$, using

\begin{align}
q=\frac{\eta \rho_*}{t_h}
\end{align}

%%AG $eta/th$ is the specific of rate of mass return to the ISM.  Ciotti et al. 1991 -- see in particular Table 2. 
Where $t_h$ is the Hubble time, and $\eta$ is an efficiency parameter--for a stellar population between 2-15 Gyr old, this number will lie in the range $\sim$0.1-1 (see Appendix~\ref{app:eta}). We adopt $\eta=1$. However, scaling $\eta$ by some factor, would simply scale the $\rho$ by the same factor and otherwise leave the solution unchanged.

% {\bf NCS: here is some added detail about the role of stellar mass
% functions in the model}

% if we consider different star formation scenarios (for example,
% continuous star formation) or different IMFs, it will change.  Once
% these free parameters are specified, however, we can answer an
% important question: do young stellar populations increase or
% decrease the SMBH feeding rate $\dot{M}$?  Clearly, $\eta(t)$ is
% larger for young stellar populations, but these stars also have high
% wind velocities that diminish the stagnation radius.  Crudely
% approximating $v_{\rm w}=200~{\rm km~s}$ for $m_{\rm max} <
% 10M_{\odot}$ and $v_{\rm w}=1000~{\rm km~s}$ for $m_{\rm max} >
% 10M_{\odot}$ (motivated by the transition from dust-driven wind loss
% on the AGB to line-driven wind loss on the main sequence), we can
% employ the relation $\dot{M} \propto \eta(t) r_{\rm
% s}^{2-\Gamma}\propto \eta(t) v_{\rm w}^{-4+2\Gamma}$ (where $\rho_*
% \propto r^{-\Gamma}$) to determine the impact of stellar ``youth''
% on SMBH feeding rates.
%
%\begin{figure}
%\includegraphics[width=\columnwidth]{NickPlot.eps}
%\caption{\label{NickPlot} SMBH feeding rates $\dot{M}=\eta(t) \times M_*(r_{\rm s})$, in arbitrary units.  The green, blue, and black curves are for galaxies with $\Gamma$ values of $0.1$, $0.5$, and $0.9$, respectively.  Solid curves represent impulsive-mode star formation, while dotted curves represent continuous-mode star formation.}
%\end{figure}
%
%%In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a function of time, for three different stellar density profiles $\Gamma = \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an influence radius $r_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
%\begin{equation}
%\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
%\end{equation}
%The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.

\section{Solutions}
\subsection{Analytics: Flow properties at $\rs$}
It is possible to derive an analytic expression for the steady-state stagnation radius, $\rs$, in terms of $v_w$ and the power law slope of the density (at $\rs$) $n$.   We summarize the main results here and place the detailed derivation in Appendix~\ref{app:rs}.

@@

\begin{align}
 &\frac{\kb T}{\mu \mp}=\gammafi \kew \label{eq:tstag}\\
 &T=6 \times 10^6 {\rm K} \, \mu \, \left({\frac{\vw}{500 \,{\rm km/s}}}\right)^2 
\end{align}

We may also write an (implicit) analytic expression for $\rs$
itself. Let  $\eta=v_{w,0}/\sigma_{\rm soi}$. Then

@@

\begin{align}
&\rs=\frac{7}{2}\frac{G \Mbh}{n \vw^2}\\
&\simeq 6 \, \pc \left(\MbhNorm\right) \left(\vwOFH\right)^{-2}
\label{eq:stag_simple}
\end{align}

@@

\begin{align}
\dot{M}\simeq
\begin{cases}
   1.5 \times 10^{-3}  \left(\Mbheight\right)^{1.9}
   \left(\vwOFH\right)^{-4} \eta \Msun \, \pyear& \text{for core galaxies} \\
   3.3 \times 10^{-3} \left(\Mbheight\right)^{1.4}
   \left(\vwOFH\right)^{-2} \eta \Msun \, \pyear  & \text{for cusp
     galaxies} 
 \end{cases}
\label{eq:mdot_analytic}
\end{align}

@@

\begin{align}
\rho(\rs)\simeq
\begin{cases}
2.5 \times 10^{-24} \Mbheight^{-0.14} \vwOFH^{-1} \eta \, {\rm g \, cm^{-3}}& \text{for   galaxies}\\
5.5 \times 10^{-24}  \Mbheight^{-0.56} \vwOFH    \eta \, {\rm g \,cm^{-3}} & \text{for cursp galaxies}
\end{cases}
\label{eq:rhors}
\end{align}

@@

%%AG: Describe conservation checks to additionally validate the results.

@@

Also, MSPs may also be dissociated in the galactic center...
%%Also low thermalization efficiency inferred in 47 Tuc by Naiman et al.

@@

Following \citealt{ShcherbakovWong+:2014a}, $E_{\rm Ia}=10^{51}$ ergs
and $R_{\rm Ia}\simeq 4 \times 10^{-14} \Msun^{-1}$ yr$^{-1}$, which
gives $\vwO\simeq 230 \, \eta^{-1/2}$ km/s (For $\eta=0.1$ this
corresponds to $\vwO\simeq$ 500 km/s.
 %Perhaps it would be best to adopt a fiducial eta=0.1--this may be more straightforward from the pov of both heating and cooling.

 As noted in \citealt{ShcherbakovWong+:2014a}, SNe energy injection
 cannot necessarily be treated as constant throughout time and space,
 since the flow time-scale may be short compared to the time between
 successive SNe Ia.

There will be some radius, $\rIa$, where the dynamical timescale of
the flow becomes comparable to the time between successive Type Ia
SNe.  If the stagnation radius ($\rs$) of the flow is outside of
$\rIa$ then SNes will provide a steady contribution to $\vwO$

@@

The stagnation radius scales linearly with $\Mbh$ (see
Equation~\ref{eq:stag_simple}), while the $\rIa$ is roughly
constant. For small $\Mbh$, $\rs$ is well inside $\rIa$ and SNes will
not contribute to the $\vwO$.  Above $\sim10^8 \Msun$ SNes will
represent an important contribution to the overall heating.
% Thus, for $\Mbh=10^6 \Msun$, we may have $\vwO\simeq 100$ km/s with
% heating mostly coming from main sequence winds. On the other hand,
% for $\Mbh \gsim 10^8 \Msun$, we may have $\vwO\simeq 500$ km/s, with
% the main contribution coming from type Ia SNe. These two regimes are
% illustrated in Figure~\ref{fig:vw_eff}.
%\end{enumerate}

@@

\subsection{Mass accretion rate}
For each of our solutions, we may calculate the accretion rate
$\dot{M}$ onto the SMBH. We show $\Mdote$ vs. $\Mbh$ as a function of
black hole mass for our three chosen values of the $\vwO$ (200 km/s,
500 km/s, and 1000 km/s). The results are shown in
Figure~\ref{fig:mdot_mass}.  Note that $\dot{M}\sim\vwO^{-2}$ for a
cusp galaxy and $\dot{M}\sim\vwO^{-4}$ for a core galaxy. Thus, the
mass accretion rate will be a sensitive function of the assumed
heating rate.
%%AG-The variation in accretion rate for w/ vw for core galaxies is comparable to that of cusp galaxies and belies the steep dependence stated above.
%%AG-Note separation of cores and cusps (at vw=1000 km/s, where there are enough core galaxies for this comparison to be made). Also, note that the spread in core galaxies is far larger.

@@

%%AG:  Also want to include a plot of x-ray luminosity (less abstract than the accretion rate).

\subsection{Comparison with observations}
\citealt{AllenDunn+:2006a} use Chandra x-ray observations to infer the
temperature and density profiles for nine nearby x-ray luminous
elliptical galaxies.  Five of the galaxies in their sample overlap
with that in \citetalias{WangMerritt:2004a}.

We may compare our results for the $T$ and $\rho$ profiles with those
inferred from observations for the overlapping galaxies.  Note that we
we have two free parameters ($\eta$ and $\vwO$ in our model and thus
we can can always match $T$ and $\rho$ at least one point.  We
generally find that $\vwO=500$ km/s (as would be expected from SNe Ia)
gives rough agreement with the observed temperatures.  Thus, we choose
$\vwO=500$ km/s.  We choose an $\eta$ so that our solution would be
consistent with the observed density profiles. The comparison is shown
in Figure~\ref{fig:allen_compare}.
%%AG-The chosen properties, in particular eta may be summarized in a table.

\begin{itemize}
\item \emph{NGC4486} %eta=0.2
  On scales of $\sim1$ kpc the density profile in
  \citealt{AllenDunn+:2006a} is shallower than in our model. Our model
  has a break in the gas density at $\sim 300$ pc--near the location
  of the Nuker break radius for this galaxy at 560 pc. There is an
  observed break in the density, but on a scale of a few kpc.

  In contrast to a naive extrapolation of the observed $T$ and $\rho$
  to smaller scales, our models have $T$ and $\rho$ monotonically
  increasing towards the galactic center. The increase in $T$ is due
  to the increased velocity dispersion (and thus increased heating)
  towards the galactic center.
\end{itemize}

\begin{figure*}
\includegraphics[width=\columnwidth]{T_compare.eps}
\includegraphics[width=\columnwidth]{dens_compare.eps}
\caption{\label{fig:allen_compare} Comparison of $n_e$ and $T$ to the \citealt{AllenDunn+:2006a}}. 
%%AG in the future we will have to be carfeful of the mean molecular weight.
\end{figure*}

@@

For each of our solutions we compare the radiative cooling rate, $C$
to the heating rate $H=q \vw^2/2$. If $H/C \gg 1$ then cooling may be
neglected. We take

@@

Where $n$ is the number density and $\Lambda(T)$ is the cooling
function. We approximate $\Lambda(T)$ as piece-wise power law (fitted
for solar metallicity--see Figure 34.1 in \cite{Draine:2011a})
%cite Draine

@@

\subsection{Angular Momentum}
\label{sec:ang}
Our models are spherically symmetric and do not include the effects of
the gas' angular momentum. However, galaxies will in general have some
net rotation, and the gas will generally have some non-zero angular
momentum. At some radius, $\rcirc$, the specific angular momentum of
the gas will equal the specific angular momentum of a circular
orbit. Inside of this radius our assumptions of no rotation and
spherical symmetry will break down. But if $\rcirc$ inner sonic point
of the flow, the region where angular momentum becomes important will
be causally disconnected from the rest of the flow, and we would be
justified in neglecting angular momentum (at least on large scales).

The inner sonic point, $\rss$ general occurs at $\simq 1\%$ of
$\rinf$. $\rcirc$ depends on the gas injection radius. The largest
injection radius is $\rs$. Let $\lrs$ be the gas specific angular
momentum at $\rs$, then

@@

 used two-dimensional kinematic data to measure the ratio of ordered
 to random motion in a sample of early type galaxies. More precisely,
 they measure

\begin{align}
\lambda_R=\frac{\left<R|V|\right>}{\lef<R\sqrt{V^2+\sigma^2}\right>}
\end{align}

Where $R$ is the distance is the distance from the galactic center, $V$ is
the mean stellar velocity, and $\sigma$ is the velocity
disperison. The brackets represent a luminosity weighted average.

$\lrs\simeq\lambdars \sigma(\rs)^2 \rs^2$ 
\section{Summary}

\appendix
\clearpage
\section{Analytic Expression for the stagnation radius}
\label{app:rs}
\input{rs}

>>>>

\documentclass[usenatbib,fleqn]{mn2e}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{outdir=./figures/}
\graphicspath{{./figures/}}
\usepackage{url}
\usepackage{aas_macros}
\usepackage{astro}
% \usepackage{natbib}

% \renewcommand{\l}{\left}
% \renewcommand{\r}{\right}

@@

\newcommand{\Menc}{M_{\rm enc}}
\newcommand{\rhostar}{\rho_*}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\Mseight}{M_{\star,8}}
\newcommand{\Mbh}[1][]{M_{\bullet#1}}
\newcommand{\Mbheight}{M_{\bullet,8}}
\newcommand{\MbheightExp}{\frac{\Mbh}{10^8 \Msun}}

@@

%% Maybe active galaxies constitute 10% need reference here?
Supermassive black holes (SMBHs) are present in the centers of most,
if not all nearby galaxies (see reviews by,
e.g. \citealt{KormendyRichstone:1995a};
\citealt{FerrareseFord:2005a}). Roughly $\sim 1\%$ of these manifest
themselves as luminous Active Galactic Nuclei. However, quiescent
black holes constitute a silent majority.

In order to understand why most galaxies appear to be inactive, it is
necessary to understand the gas environment of black holes, as the gas
density around the black hole sets its accretion rate.  The accretion
rate also has implications for rate of black hole growth.
% AG:Perhaps need more detail here.

Constraints on the SMBH accretion rates at the low end of the SMBH
mass function also have implications for the black occupation
fraction.  For example, \citealt{MillerGallo+:2014a} used an observed
relation between $\Mbh$ and nuclear x-ray luminosity $L_x$ to
constrain the BH occupation fraction.  However, they extrapolate
(assuming a single power law) their observed relationship below their
detection threshold.  However, it is possible that the $L_x$-$\Mbh$
relationship steepens at smaller $\Mbh$, and thus the black hole
occupation fraction could be underestimated at low $\Mbh$.
% AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

@@

However, this approach has limitations. Due to resolution issues it is
almost never possible to observationally determine the gas density at
the black sphere of influence, and such studies are forced to
extrapolate observed gas temperatures and densities to smaller scales.
In fact, the temperature profile is assumed to be flat inside of the
Bondi radius.  In reality, the cusp in the velocity dispersion near
the SMBH, should cause a cusp in the gas temperature profile (assuming
the kinetic energy of stellar winds is efficiently thermalized in
shocks).
%% AG-give sense of scale.

@@

We compare our calculated gas profiles to those inferred from Chandra
X-ray observations by \citealt{AllenDunn+:2006a} and
\citealt{RussellMcNamara+:2013a}. Additionally, we chack how well our
results are able to reproduced relationships between $L_x$ and $\Mbh$
in \citealt{MillerGallo+:2014a}. In performing these comparisons we
hope to determine how well the gas environments of quiescent galaxies
may be understood using the relatively well understood physics of
stellar winds and supernova heating.
%% AG--say something about x-ray observations.

@@

This is a broken power law which transitions from an inner power law
slope, $\gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  The corresponding (spherically symmetric) stellar
density, $\rhostar\sim r^{-1-\gamma}$ for $r \ll \rb$ and
$\rhostar\sim r^{-1-\beta}$ for $r \ll \rb$.  $\gamma\simeq1$ for cusp
galaxies, while $\gamma\simeq0$ for core galaxies.
% AG:Is there a systematic dependence of beta on galaxy type$.

@@

\section{Model}
\label{sec:model}
We use the same model as \citealt{Quataert:2004a} to find a steady
state gas density near the SMBH for each of the galaxies within our
sample.  Gas is supplied by winds from the stellar population in the
galactic center. The equation we solve are

\begin{align}
  &\drhodt+\frac{1}{r^2}\dxdy{\rho r^2 v}{r}=q\\
  &\rho \frac{dv}{dt}=-\dpdr-\rho \dphidr-q v\\
  &\rho T \frac{ds}{dt}=q\left[\ke+\kew+\gammaf \cs \right].
\end{align}

Note that we assume spherical symmetry, an ideal gas equation of
state, and rate of mass injection $q$. $q$ is related to related to
the stellar density $\rhostar$ via

\begin{align}
  q=\frac{\eta \rho_*}{\th}.
\end{align}

%% AG $eta/th$ is the specific of rate of mass return to the ISM.  Ciotti et al. 1991 -- see in particular Table 2. 
Where $t_h$ is the Hubble time, and $\eta$ is an efficiency parameter--for a stellar population between 2-15 Gyr old, this number will lie in the range $\sim$0.1-1 (see Appendix~\ref{app:eta}). We adopt $\eta=1$. However, scaling $\eta$ by some factor, would simply scale the $\rho$ by the same factor and otherwise leave the solution unchanged.

@@

\section{Solutions}
\subsection{Analytics: Flow properties at $\rs$}
It is possible to derive an analytic expression for the steady-state
stagnation radius, $\rs$, in terms of $v_w$ and the power law slope of
the density (at $\rs$) $n$.  We parameterize our results in terms of
$\vwOFH=\vwOFHexp$, $\Mbheight=\MbheightExp$ and
$\xi=\left(1+\frac{0.14 \Mbheight^{0.43}}{\vwOFH}\right)^{1/2}$ . $xi$
encapsulates the contribution of $\sigma$ in the heating.  We
summarize the main results here and place the detailed derivation in
Appendix~\ref{app:rs}.

@@

\begin{align}
  &\frac{\kb T}{\mu \mp}=\gammafi \kew \label{eq:tstag}\\
  &T=6 \times 10^6 {\rm K} \,  \,\ \vwO500^2 \xi^2 \mu
\end{align}

%%Need this to be consistent with our new vw convention.
We may also write an (implicit) analytic expression for $\rs$
itself. Let  $\eta=v_{w,0}/\sigma_{\rm soi}$. Then

@@

\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n \vw^2}\\
  &\simeq 6 \, \pc \,\, \Mbheight \left(\vwOFH\right)^{-2}
  \label{eq:stag_simple}
\end{align}

@@

\begin{align}
  \dot{M}\simeq
  \begin{cases}
    1.6 \times 10^{-3}  \left(\Mbheight\right)^{1.9}
    \left(\vwOFH\right)^{-4} \xi^{-4} \eta \Msun \, \pyear& \text{for cores} \\
    3.4 \times 10^{-3} \left(\Mbheight\right)^{1.4}
    \left(\vwOFH\right)^{-2} \xi^{-2} \eta \Msun \, \pyear  & \text{for cusps} 
  \end{cases}
  \label{eq:mdot_analytic}
\end{align}

@@

\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    2.6 \times 10^{-24} \Mbheight^{-0.14} \vwOFH^{-1} \xi^{-1} \eta \, {\rm g \, cm^{-3}}& \text{for cores}\\
    5.6 \times 10^{-24}  \Mbheight^{-0.56} \vwOFH  \xi^{-1}  \eta \, {\rm g \,cm^{-3}} & \text{for cusps}
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

%% AG: Describe conservation checks to additionally validate the results.

@@

  Also, MSPs may also be dissociated in the galactic center...
  %% Also low thermalization efficiency inferred in 47 Tuc by Naiman et al.

@@

  % Following \citealt{ShcherbakovWong+:2014a}, $E_{\rm Ia}=10^{51}$ ergs
  % and $R_{\rm Ia}\simeq 4 \times 10^{-14} \Msun^{-1}$ yr$^{-1}$, which
  % gives $\vwO\simeq 230 \, \eta^{-1/2}$ km/s (For $\eta=0.1$ this
  % corresponds to $\vwO\simeq$ 500 km/s.
  % Perhaps it would be best to adopt a fiducial eta=0.1--this may be more straightforward from the pov of both heating and cooling.

  As noted in \citealt{ShcherbakovWong+:2014a}, SNe energy injection
  cannot necessarily be treated as constant throughout time and
  space. This would only make sense if many SNes are exploding on the
  dynamical timescale of the flow.

  To check this quantitatively we may compare the radius where the
  dynamical timescale becomes comparable to the time between successive
  Ias, $\rIa$, to the stagnation radius, $\rs$. If $\rIa>>\rs$, SNes
  will provide a steady contribution to the $\vwO$.

@@

  The stagnation radius scales linearly with $\Mbh$ (see
  Equation~\ref{eq:stag_simple}), while the $\rIa$ is roughly
  constant. For small $\Mbh$, $\rs$ is well inside $\rIa$ and SNes will
  not contribute to the $\vwO$.  Above $\sim10^8 \Msun$ SNes will
  represent an important contribution to the overall heating.
  % Thus, for $\Mbh=10^6 \Msun$, we may have $\vwO\simeq 100$ km/s with
  % heating mostly coming from main sequence winds. On the other hand,
  % for $\Mbh \gsim 10^8 \Msun$, we may have $\vwO\simeq 500$ km/s, with
  % the main contribution coming from type Ia SNe. These two regimes are
  % illustrated in Figure~\ref{fig:vw_eff}.
  % \end{enumerate}

@@

\subsection{Mass accretion rate}
For each of our solutions, we may calculate the accretion rate
$\dot{M}$ onto the SMBH. We show $\Mdote$ vs. $\Mbh$ as a function of
black hole mass for our three chosen values of the $\vwO$ (200 km/s,
500 km/s, and 1000 km/s). The results are shown in
Figure~\ref{fig:mdot_mass}.  Note that $\dot{M}\sim\vwO^{-2}$ for a
cusp galaxy and $\dot{M}\sim\vwO^{-4}$ for a core galaxy. Thus, the
mass accretion rate will be a sensitive function of the assumed
heating rate.
%% AG-The variation in accretion rate for w/ vw for core galaxies is comparable to that of cusp galaxies and belies the steep dependence stated above.
%% AG-Note separation of cores and cusps (at vw=1000 km/s, where there are enough core galaxies for this comparison to be made). Also, note that the spread in core galaxies is far larger.

@@

%% AG:  Also want to include a plot of x-ray luminosity (less abstract than the accretion rate).

\subsection{Comparison with observations}
\subsubsection{Comparison with observed gas temperatures and densities}
\citealt{AllenDunn+:2006a} use Chandra x-ray observations to infer the
temperature and density profiles for nine nearby x-ray luminous
elliptical galaxies.  \citealt{RussellMcNamara+:2013a} performed a
similar analysis with a somewhat enlarged sample.
% At some point it would be useful to describe the differences between
% these two sets of observations \citealt{RussellMcNamara+:2013a}
% tries several different extrapolation schemes for the gas density,
% and finds that that the uncertainty in extrapolating the density
% profile is the dominant source of uncertainty in measuring the
% accretion rate. Find total number which overlaps once the Russell
% galaxies are included...
Five of the galaxies in their sample overlap with that in \citetalias{WangMerritt:2004a}.

We may compare our results for the $T$ and $\rho$ profiles with those
inferred from observations for the overlapping galaxies.  Note that we
we have two free parameters ($\eta$ and $\vwO$) in our
model. Therefore, we can always match the observed $T$ and $\rho$
profiles for at least one point.  We generally find that $\vwO=500$
km/s (as would be expected from SNe Ia) gives rough agreement with the
observed temperatures. 
% Thus, we choose $\vwO=500$ km/s.  We choose an
% $\eta$ so that our solution would be consistent with the observed
% density profiles. The comparison is shown in
Figure~\ref{fig:allen_compare}.
%% AG-The chosen properties, in particular eta may be summarized in a table.
%% AG-Mention rising temperature profiles on large scales.

\begin{itemize}
% \item \emph{NGC4486} %eta=0.2
%   On scales of $\sim1$ kpc the density profile in
%   \citealt{AllenDunn+:2006a} is shallower than in our model. Our model
%   has a break in the gas density at $\sim 300$ pc--near the location
%   of the Nuker break radius for this galaxy at 560 pc. There is an
%   observed break in the density, but on a scale of a few kpc.

%   The temperature is assumed to be flat from the resolution limit
%   inwards in \citealt{AllenDunn+:2006a}. However, in our models $T$
%   increases monotonically towards the galactic center. This increase
%   is due to the increased velocity dispersion (and thus increased
%   heating) towards the galactic center.

\item \emph{NGC5813} %eta=0.1
  The observed density profile \citealt{RussellMcNamara+:2013a} is
  shallower than our calculated density profile, which has a sharp
  break at $\sim$0.1 kpc, the location of the break radius in the
  Nuker profile (this is the radius where the stellar light profile
  transitions from a shallower inner power law slope to a steeper
  outer power law slope.

  For this galaxy we found that we could reproduce the observed
  temperature on large scales with $\vwO=500$ km/s. The increased
  velocity disperision towards the galactic center results in a sharp
  increase in $T$ at small radii, which is not captured in the
  observations due to resolution issues.
\end{itemize}

%It would be nice to have error bars for the observational points in this plot...
\begin{figure*}
  \includegraphics[width=\columnwidth]{NGC5813_T.eps}
  \includegraphics[width=\columnwidth]{NGC5813_rho.eps}
  \caption{\label{fig:allen_compare} Comparison of $n_e$ and $T$ to
    the profiles in \citealt{AllenDunn+:2006a} and \citealt{ RussellMcNamara+:2013a}}.
  %% AG in the future we will have to be carfeful of the mean molecular weight.
\end{figure*}

@@

To assess the importance of radiative cooling for our solutions we
compare the radiative cooling rate, $C$ to the heating rate $H=q
\vw^2/2$. If $H/C \gg 1$ then cooling may be neglected. We take

@@

%% at low densities there might be some sort of non-CIE cooling operating.
Where $n$ is the number density and $\Lambda(T)$ is the cooling
function. We approximate $\Lambda(T)$ as piece-wise power law (fitted
for solar metallicity--see Figure 34.1 in \cite{Draine:2011a})

@@

  \subsection{Angular Momentum}
  \label{sec:ang}
  Our models are spherically symmetric and do not include the effects of
  the gas' angular momentum. However, galaxies will in general have some
  net rotation, and the gas will generally have some non-zero angular
  momentum. At some radius, $\rcirc$, the specific angular momentum of
  the gas will equal the specific angular momentum of a circular
  orbit. Inside of this radius our assumptions of no rotation and
  spherical symmetry will break down. But if $\rcirc$ is inside of the
  inner sonic point of the flow, the region where angular momentum
  becomes important will be causally disconnected from the rest of the
  flow, and we would be justified in neglecting angular momentum (at
  least on large scales).

  $\rcirc$ depends on the gas injection radius. The largest
  injection radius is $\rs$. Let $\lrs$ be the gas specific angular
  momentum at $\rs$, then

@@

  \citealt{EmsellemCappellari+:2007a} used two-dimensional kinematic
  data to measure the ratio of ordered to random motion in a sample of
  early type galaxies. More precisely, they measure

  \begin{align}
    \lambda_R=\frac{\left<R|V|\right>}{\left<R\sqrt{V^2+\sigma^2}\right>}
  \end{align}

  Where $R$ is the distance is the distance from the galactic center, $V$ is
  the mean stellar velocity, and $\sigma$ is the velocity
  dispersion. The brackets represent a luminosity weighted average.

  We may rewrite the specific angular momentum at $\rs$ in terms of $\lambda_R$.

  \begin{align}
    \lrs\simeq\lambdars \sigma(\rs)^2 \rs^2
  \end{align}

  Then

  \begin{align}
    \rcirc&\simeq\lambdars^2 \frac{\sigma(\rs)^2}{G \Mbh} \rs^2\\
    &\simeq\lambdars^2 \frac{\rs^2}{\rsoi} \lsim \lambdars^2 \rsoi
  \end{align}

  The inner sonic point, $\rss$ general occurs at $\simeq 1\%$ of
  $\rsoi$. %%AG-Need to explain this somehow.
  So if $\lambdars\lsim 0.1$, $\rcirc<\rss$. In Figure 2 of
  \citealt{EmsellemCappellari+:2007a} $\lambda_R$ is plotted as function
  of radius for a sample of early type galaxies. On scales of $\sim 0.1
  R_e$ $\lambda_R$ is either less than 0.1 or rapidly declining for most
  of their sample %%need to be clearer on rapidly declining also compare
  %% this scale to the sphere of influence scale.
  This suggests that angular momentum should generally be unimportant.

  \section{Summary}

  \clearpage
  \appendix
  \section{Analytic Expression for the stagnation radius}
  \label{app:rs}
  \input{rs}
